---
title: "Testing DBNTree on synthetic data"
author: "David Quesada"
date: "22/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Testing mtDBN on synthetic data

This markdown document will follow the experimental procedures of applying the mtDBN model to the synthetic dataset generated by the 'dt_generation_notebook.Rmd' markdown. The objective of these experiments is to show the better performance of mtDBN models over regular DBN models when dealing with non-linear relationships and non-stationary time-series. 

First, the synthetic dataset will be split in 67% train and 33% test. With that train dataset, both the mtDBN and regular DBN models will be trained and then they will perform forecasting on the test cycles. 

## Dataset split

To begin the experiment, we will split the dataset into train and test, where 67% will be training data and 33% will be test. This split will be done by selection whole cycles from the original dataset. Once the split is performed, the datasets have to be folded to size = 2 and the instances that have overlapping values from different cycles have to be removed.

```{r init}
library(dbnR)
library(mtDBN)
library(data.table)
```

```{r prepare datasets}
dt <- fread("../dataset/dt_cycles.csv")
set.seed(42)
train_p <- 0.67

vars_noise <- c("rho_1", "C_p1", "C_in", "vol", "C_ain") # Add some noise to avoid singular matrix errors later
dt[, eval(vars_noise) := .SD + rnorm(100000, 0, 0.5), .SDcols = vars_noise]

ids <- dt[, unique(cyc)]
train_ids <- sample(ids, round(length(ids) * train_p), replace = F)

dt_train <- dt[cyc %in% train_ids]
dt_test <- dt[!(cyc %in% train_ids)]

size <- 2
id_var <- "cyc"

f_dt_train <- dbnR::filtered_fold_dt(dt_train, size, id_var)
f_dt_test <- dbnR::filtered_fold_dt(dt_test, size, id_var)
dt_train[, (id_var) := NULL]
dt_test[, (id_var) := NULL]
```

Once the dataset is prepared for training, we are ready to fit both the mtDBN and regular DBN models and see how well they perform on synthetic data

```{r fit network}
mode <- 2
set.seed(42)
mv <- F
homogen <- T

if(mv){
  obj_vars <- c("T_1","T_2", "S_c", "C_a", "C_p1", "rho_1")
}else{
  obj_vars <- "T_1"
}

if(mode == 1){
  model_net <- dbnR::learn_dbn_struc(dt_train, size, method = "dmmhc", f_dt = f_dt_train)
  model_fit <- dbnR::fit_dbn_params(model_net, f_dt_train)
  model <- mtDBN::mtDBN$new()
  model$fit_model(dt_train, size, method = "dmmhc", obj_var = obj_vars, mv = mv, homogen = homogen, min_ind = 300, max_depth = 6, f_dt = f_dt_train, prune_val = 0.045)
} else{
  # Single dbn model
  print("Single model training.")
  tmp <- Sys.time()
  model_net <- dbnR::learn_dbn_struc(dt_train, size, method = "psoho", f_dt = f_dt_train, n_it = 20, n_ind = 50, gb_cte = 0.5, lb_cte = 0.5, r_probs = c(-0.5, 1.5), v_probs = c(10,65,25))
  model_fit <- dbnR::fit_dbn_params(model_net, f_dt_train, replace.unidentifiable = T)
  print(tmp - Sys.time())
  
  print("----------------------")
  
  # mtDBN model
  print("mtDBN training.")
  tmp <- Sys.time()
  model <- mtDBN::mtDBN$new()
  model$fit_model(dt_train, size, method = "psoho", obj_var = obj_vars, mv = mv, homogen = homogen, min_ind = 300, max_depth = 6, f_dt = f_dt_train, n_it = 20, n_ind = 50, gb_cte = 0.5, lb_cte = 0.5, r_probs = c(-0.5, 1.5), v_probs = c(10,65,25), prune_val = 0.045)
  print(tmp - Sys.time())
  print("Model training finish.")
  
} 

```

```{r forecasting, eval=FALSE}
ini <- 299
len <- 94

obj_vars <- c("T_1_t_0", "C_a_t_0")
ev_vars <- c("C_a_t_0")
ev_vars <- NULL

print("Forecasting time for single net: ")
res_net <- dbnR::forecast_ts(f_dt_test, model_fit, size = size, obj_vars = obj_vars, ini = ini, len = len, prov_ev = ev_vars)

print("----------------------")

print("Forecasting time for mtDBN")
res <- model$forecast_ts(f_dt_test, obj_vars, ini = ini, len = len, prov_ev = ev_vars)

```
